---
title: "DATA698 Project - Vehicle Loan Default Prediction"
author: "Chinedu Onyeka"
date: "12/03/2023"
output:
  html_document:
    includes:
      in_header: header.html
    css: ./lab.css
    highlight: pygments
    theme: cerulean
    toc: true
    toc_float: true
    code_folding: "hide"
  pdf_document: default
  prettydoc::html_pretty:
    theme: cayman
    highlight: github
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
#library(tidyverse)
```


<center> <h2> 1. INTRODUCTION </h2> </center> 
<h3> 1.1 Research Problem </h3> 
<h3> 1.2 Abstract </h3> 
<center> <h2> 2. LITERATURE REVIEW </h2> </center> 

<center> <h2> 3. METHODOLOGY </h2> </center> 
<h3> 3.1 DATASET </h3> 
<h3> 3.2 Data Pre-processing </h3> 
<h4> 3.2.1 Data Cleaning </h4> </center> 
<h4> 3.2.2 Feature Selection and pre-processing </h4>

<center> <h2> 4. INTRODUCTION TO MODELS </h2> </center> 
<h3> 4.1 Predictive Analytics Overview </h3>
<h3> 4.2 Scoring the Model and understanding the scoring metrics </h3>
<h3> 4.3 Logistic Regression </h3>
<h3> 4.4 Tree-Based Methods </h3> 
<h4> 4.4.1 Random Forest </h4> 
<h4> 4.4.2 Gradient Boosting </h4>

<center> <h2> 5. ANALYSIS, TESTING AND RESULTS </h2> </center> 
<h3> 5.1 Analysis </h3>
<h3> 5.2 Testing </h3>
<h3> 5.3 Results </h3>

<center> <h2> 6. CONCLUSION </h2> </center> 



